Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-01-04 src/ms.tex
minor edits

\begin{document}
\title{Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}

@@

{\color{red} The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling. The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling. Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling. In this picture, surface spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability. After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there. A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge.}

@@

    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pile-up. For example, a Sun-like star may spend several Gyr evolving across the ridge. Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim 5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.

>>>>

\begin{document}
\title{Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}

@@

{\color{red} The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling. The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling. Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling. In this picture, surface spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability. After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there. A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge. We note that this is a speculative scenario and the underlying physics are not known.}

@@

    \item The existence of the long-period pile-up limits the utility of gyrochronology for the hottest stars with convective envelopes, as stellar spin-down appears to stall on the pile-up. For example, a Sun-like star may spend several Gyr evolving through the long-period pile-up. Authors using gyrochronology as a means of age-dating a field dwarf star with \teff~$\gtrsim 5800~K$ (and possibly cooler \teff as well) should take care to assess whether that star resides on the long-period edge, in which case the uncertainty on the age may be larger than current gyrochronology calibrations imply.
